Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-21 Master Thesis/synopsis.tex
Add half-reference for 2.2-pre
Fix typo

@@

\paragraph{} Let $K$ be a (not necessarily finitely generated) field of characteristic $0$ and dimension $2$.

@@

\begin{lemma}\phantomsection\label{4.5}
\begin{enumerate}[(i)]
\item There exists a finite subset $S\subset X_0$ such that:
\[ \operatorname{Pic}(X\setminus S)=0=\operatorname{Pic}(X'\setminus S') \]
Such a set $S$ is called \textit{admissible}. If $V\subset X$ is an open subscheme, then one can choose $S$ in $V_0$.
\item Let $S$ be admissible. For all $n\in\mathbb{N}$ the following canonical maps are injective:
\[ K^S/n \to K^\times /n,\quad K'^{S'}/n\to K'^\times /n \]
$\varphi_n$ induces isomorphisms $\varphi_n: K'^{S'}/n\to K^S/n$.
\end{enumerate}
\end{lemma}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$. It follows $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i),\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}\subset U_0$, the map $r_{S,T}$ is injective. For the second assertion, note the following isomorphism:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \Big(\prod_{P\in T}k(P)^\times \Big)\Big/k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}

>>>>

\begin{proof}
See Valued Fields 4.1.4.
\end{proof}

@@

Let $K$ be a (not necessarily finitely generated) field over $\mathbb{Q}$ and dimension $2$.

@@

\begin{lemma}\phantomsection\label{4.5}
\begin{enumerate}[(i)]
\item There exists a finite subset $S\subset X_0$ such that:
\[ \operatorname{Pic}(X\setminus S)=0=\operatorname{Pic}(X'\setminus S') \]
Such a set $S$ is called \textit{admissible}. If $V\subset X$ is an open subscheme, then we can choose $S$ in $V_0$.
\item Let $S$ be admissible. For all $n\in\mathbb{N}$ the following canonical maps are injective:
\[ K^S/n \to K^\times /n,\quad K'^{S'}/n\to K'^\times /n \]
$\varphi_n$ induces isomorphisms $\varphi_n: K'^{S'}/n\to K^S/n$.
\end{enumerate}
\end{lemma}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
Clearly, (i) and (ii) hold for almost all $a\in k^\times$. We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$ and $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i)^\times,\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}\subset U_0$, the map $r_{S,T}$ is injective. For the second assertion, note the following isomorphism:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \Big(\prod_{P\in T}k(P)^\times \Big)\Big/k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}
